``FISTA'' in Banach spaces with adaptive discretisations

### 2022-03-14 response.tex

\section{Detailed response to first reviewer}
\review{If I had one more comment it would be that I find it somewhat odd the value of the paper's algorithm is presented ``for use with un-attained minima'' as the key theoretical novelty. As a point of critique: what does it mean to solve an optimization problem that doesn't have a solution? Do the solutions for the $U^n$-problems become meaningless asymptotically? Have the authors observed any odd behavior for large values of $n$? On the other hand, I think the authors sell themselves a bit short with the statement ``key theoretical novelty'' as the analysis could be useful for solving optimal control and PDE-constrained optimization problems with FISTA or related algorithms, where the interplay of mesh-size and algorithm convergence should always be taken into account.}

Thank you for the constructive feedback. That paragraph in the introduction has been re-worded to take into account some of your points. In answer to your questions, in an abstract sense one can always `attain' minima by adding points to (i.e. `completing') the space. As you say, the challenge is to have some regularity guarantees so that the iterates themselves are not `meaningless', but the meaning of meaningless depends on the topology.

A simple 1D example, if $E(u) = e^{-u}$ for $u\geq0$, then numerically I get $u_n\approx 4\log(n)$. Clearly $|u_n-\infty|\not\to0$, but in a `weaker sense', certainly $u_n\to\infty$. This seems to be representative of the general behaviour. In the Lasso example, the convergence of iterates is meaningless in both $L^2$ and $L^1$, but can be understood in the space of measures. In this sense, we have never observed anything unexpected for large $n$, the iterates do converge to tall/thin Dirac-like functions. Unfortunately this behaviour can't be understood from the values of $a_E$ and $a_U$ alone.

@@

	You're absolutely right, Theorem 4 was not consistent with the lemmas in the appendix. We have kept the statement of Theorem 4 the same other than to remove the special case and remove the implicit assumption that $\tilde{\F U}^0$ was finite dimensional (previously part of Definition 2). We have also added a new lemma (Lemma 7) to improve clarity, it serves as a midpoint between Theorem 4 and Appendix B. The text in Appendix B also makes reference to Lemma 7 as part of the proof of Theorem 4.

	\item \review{(15): $A$ has to be continuous on $H$ as well.}

	Old (15) is now (18). The assumptions on $A$, including continuity on $\F H$, are now just above (18).

	\begin{quote}
		\edit{}{Let $A\colon\F{U}\cap\F{H}\to\F R^m$ be a linear operator represented by the kernels $\psi_j\in\F{H}$ such that 
		\begin{equation*}\label{eq: kernels of A}
			\forall u\in\F{U}\cap\F{H},\ j=1,\ldots,m,\qquad (Au)_j = \langle \psi_j, u\rangle.
		\end{equation*}
		In the continuous case we will assume the additional smoothness $\psi_j\in C^1([0,1]^d)$. In Section~6.5 we will formally define and estimate several operator semi-norms for $A$ of this form, for example Lemma~8 confirms that $A$ is continuous on $\F{H}$ (without loss of generality $\|A\|\leq 1$).} 
	\end{quote}

@@

	\item \review{(25) What is $\partial_n E$?}

	Old (25) is now (27). This refers to the derivative of $E$ on the subspace $\F U^n$, this is now clarified immediately above it in the text.
	\begin{quote}
		\edit{}{We consider the discretised function $E|_{\F{U}^n}\colon \F{U}^n\to\F R$ and its subdifferential $\partial_nE(\cdot) = \Pi_n\partial E(\cdot)$ on $\F{U}^n$.} 
	\end{quote}

	\item \review{Inequalities (61),(62) do not make sense: the quantities on the left and right-hand sides are equal.}

	Please can the reviewer clarify this point? The old (61-62) are now (60-61) which are equalities. There was a pair of inequalities (59-60), now (58-59), but they do not seem to fit the comment. They are just statements of H\"older's inequality.

### 2022-03-14 paper.tex
New fig and final edits

	\section{Definitions and notation}\label{sec: prelims}
	We consider optimisation of \eqref{eq: min F = fD+fC} over a Hilbert space $(\F H,\IP\cdot\cdot,\norm\cdot)$ with the translated energy
	\begin{equation}
		\op{E}_0\colon\F H\to \R,\qquad \op{E}_0(\var0) \coloneqq \op{E}(\var0)-\Emin[\tilde{\var0}].
	\end{equation}

@@

	\begin{algorithm}\caption{Refining subset FISTA}\label{alg: refining FISTA}
		\centering
		\begin{algorithmic}[1]
			\State Choose $(\F{U}^n)_{n\in\F N}$, $\var0_0\in \F{U}^0$ and some FISTA stepsize choice $(t_n)_{n\in\F N}$
			\State $\var1_0\gets \var0_0, n\gets0$
			\Repeat
			\State $\bar{\var0}_n \gets (1-\tfrac{1}{t_n})\var0_n + \tfrac{1}{t_n}\var1_n$
			\State $\displaystyle \var0_{n+1} \gets \argmin_{\var0\in \F{U}^{n+1}} \tfrac12\norm{\var0-\bar{\var0}_n+\nabla \op{f}(\bar{\var0}_n)}^2+\op{g}(\var0)$ \Comment{Only modification, $\F{U}^{n+1}\subset \F{U}$}
			\State $\var1_{n+1} \gets (1-t_n)\var0_n + t_n\var0_{n+1}$
			\State $n\gets n+1$
			\Until{a stopping criterion is met}
		\end{algorithmic}
	\end{algorithm}

	\section{General proof recipe}\label{sec: recipe}
	In this section we give an intuitive outline of the full proof for convergence of Algorithm~\ref{alg: refining FISTA} before giving formal theorems and proofs in the next section. First we recall the classical FISTA convergence guarantee given by \cite[Thm 3.1]{Chambolle2015}\edit{:}{; if there exists $\var0^*\in \argmin_{\var0\in\F H}\op{E}(\var0)$, then}
	\begin{equation}\label{eq: classical FISTA}
		t_N^2\op{E}_0(\var0_N) + \sum^{N-1}_{n=1}\rho_n\op{E}_0(\var0_n) + \tfrac12\norm{\var1_N-\var0^*}^2 \leq \tfrac12\norm{\var0_0-\var0^*}^2
	\end{equation}
	for \edit{some $\var0^*\in \argmin_{\var0\in\F H}\op{E}(\var0)$, and}{any} FISTA stepsize choice $t_N\simeq N$ such that $\rho_n\geq0$.

@@

	\paragraph{Step 2: Quantifying the scaling properties}
	\begin{adaptiveStep}\label{step: step 2 smoothness}
		To show that the extra terms in \eqref{eq: mini FISTA convergence} are small, we need to quantify the approximation properties of $\F{U}^n$. \edit{The idea is that $\F{U}^n$ are locally compact, so that minimisers are achieved, but `growing' so that $\min_{\var0\in\F{U}^n}\op{E}(\var0) \to \inf_{\var0\in\F H}\op{E}(\var0)$.  A canonical example might be $\F{U}^n=\{\var0\in\F H\st \norm{\var0}\leq n\}$.}{The idea is that there is a sequence $\var2_n\in\F{U}^n$, $n\in\F N$ such that $\norm{\var2_n}$ grows slowly and $\op{E}_0(\var2_n)$ decreases quickly.} To quantify \edit{the}{this} balance\edit{ between a growing Hilbert norm and improving estimate of $\op{E}$}{}, we introduce a secondary sequence $n_0 < n_1 < \ldots$ and constants $\aU,\aE\geq1$ such that for each $k\in\F N$
		\begin{equation}\label{eq: first regularity}
			n\leq n_k \implies \norm{\var2_n} \lesssim \aU^k, \qquad n\geq n_k \implies \op{E}_0(\var2_n) \lesssim \aE^{-k}.
		\end{equation}
		\edit{The expanding balls example would become}{A canonical example would be} $\F{U}^n=\{\var0\in\F H\st \norm{\var0}\leq \aU^k\}$ for $n\in[n_k,n_{k+1})$, then $\aE$ reflects the smoothness of $\op{E}$. The choice of exponential scaling \edit{is more}{is introduced to improve stability of Algorithm~\ref{alg: refining FISTA}. It is} natural if we consider the $\F{U}^n$ to be the subspace of functions discretised on a uniform mesh. If that mesh is sequentially refined, then the resolution of the mesh will be of order $h^k$ after $k$ refinements and for some $h<1$. The integer $n_k$ is then the time at which the mesh has refined $k$ times. \edit{The value $\aU$ reflects the smoothness of the chosen basis, and $\aE$ is a measure of efficiency with respect to the structure of $\op{E}$. If an efficient basis is used, then $\aE$ is large.}{} The trade-off between $\aE$ and $\aU$ dictates the final convergence rate of the algorithm. If $\aU>1$, then we cannot guarantee the original $n^{-2}$ rate of convergence.
	\end{adaptiveStep}

@@

		Spatial adaptivity is robust, so long as the scaling properties of Step~\ref{step: step 2 smoothness} are satisfied. The only other constraint is to ensure that the partial telescoping of Step~\ref{step: step 3 convergence bound} still holds.%
		}{%
		The main challenge for adaptive choice of $\F{U}^n$ is to guarantee \eqref{eq: first regularity} from Step~\ref{step: step 2 smoothness} using a posteriori estimates. Combined with the partial telescoping requirement in Step~\ref{step: step 3 convergence bound}, a natural choice is $\var2_n=\var0_{n_k-1}$ for $n\in[n_k,n_{k+1})$, i.e. the value of $n_k$ is chosen to be $n+1$ once the iterate $\var0_n$ is observed.
		}
		Theorem~\ref{thm: stronger exponential FISTA convergence} shows that a sufficient condition is
		$$\var0_{n_k-1}\in\F{U}^{n_k}\cap\F{U}^{n_k+1}\cap\ldots\cap\F{U}^{n_{k+1}-1}\edit{}{, \qquad \norm{\var0_{n_k-1}} \lesssim \aU^k, \quad\text{and}\quad \op{E}_0(\var0_{n_k-1}) \lesssim \aE^{-k}}.$$
		Convergence is most stable if the approximation spaces $\F{U}^n$ satisfy a monotone inclusion, breaking the monotonicity requires more care.
		\edit{Adaptively choosing the time of refinement is more challenging due to the non-descent property of FISTA. The idea is that $\F{U}^n$ should be refined as soon as the value $\op{E}(\var0_n)$ is sufficiently close to $\op{E}(\var2_n)$. This is analysed in Theorem~\ref{thm: stronger exponential FISTA convergence} (sufficient conditions in Lemma~\ref{thm: practical refinement criteria}) which shows convergence results of the form}{The only non-trivial property to verify is the energy gap  $\op{E}_0(\var0_n) = \op{E}(\var0_n)-\Emin$. Lemma~\ref{thm: practical refinement criteria} proposes some sufficient conditions to guarantee the same overall rate of convergence as in Step~\ref{step: step 5 sufficiently slow},}
		$$ \min_{n\leq N}\op{E}_0(\var0_n) \lesssim \frac{N^{2\kappa}}{N^2}$$
		for all $N\in\F N$, with the same $\kappa\in[0,1)$ from Step~\ref{step: step 5 sufficiently slow}. The penalty for accelerating the change of discretisation is a potential loss of stability or monotonicity in $\op{E}(\var0_n)$, although this behaviour has not been seen in numerical experiments.
	\end{adaptiveStep}

@@

	\subsubsection{Single iterations}
	We first wish to understand a single iteration of Algorithm~\ref{alg: refining FISTA}. This is done through the following two lemmas.
	\begin{lemma}[equivalent to {\cite[Lemma 3.1]{Chambolle2015}}]\label{thm: descent lemma}
		\Copy{thm: descent lemma}{
			Suppose $\nabla \op{f}$ is 1-Lipschitz, for any $\bar{\var0}\in \F H$ define
			$$ \var0 \coloneqq \argmin_{\var0\in \F{U}^{n}} \tfrac12\norm{\var0-\bar{\var0}+\nabla \op{f}(\bar{\var0})}^2+\op{g}(\var0).$$
			Then, for all $\var2\in\F{U}^n$, we have
			$$\op{E}(\var0) + \tfrac12\norm{\var0-\var2}^2 \leq \op{E}(\var2) + \tfrac12\norm{\bar{\var0}-\var2}^2.$$

@@

	\begin{lemma}[{\cite[(17)]{Chambolle2015}, \cite[Lemma 4.1]{Beck2009}}]\label{thm: one step FISTA}
		\Copy{thm: one step FISTA}{
			Let $\var2_n\in\F{U}^n$ be chosen arbitrarily and $\var0_n$/$\var1_n$ be generated by Algorithm~\ref{alg: refining FISTA} for all $n\in\F N$.
			For all $n\in\F N$, it holds that}
		\begin{equation}\Copy{thm:eq: one step FISTA}{
			t_{n}^2(\op{E}(\var0_{n}) - \op{E}(\var2_n)) - (t_{n}^2-t_{n})(\op{E}(\var0_{n-1})-\op{E}(\var2_n)) \leq \tfrac1{2}\left[\norm{\var1_{n-1}}^2-\norm{\var1_{n}}^2\right] + \IP{\var1_{n}-\var1_{n-1}}{\var2_n}.}\label{eq: one-step FISTA}
		\end{equation}
	\end{lemma}
	\noindent The proof is given in Theorem~\ref{app:thm: one step FISTA} and is a result of the convexity of $\op{E}$ and $\F{U}_n$ for a well chosen $\var2$ in Lemma~\ref{thm: descent lemma}.

@@

	\begin{theorem}[analogous to {\cite[Thm 3.2]{Chambolle2015}, \cite[Thm 4.1]{Beck2009}}]\label{thm: mini FISTA convergence}
		\Copy{thm: mini FISTA convergence}{
			Fix a sequence of subsets $(\F{U}^n)_{n\in\F N}$ satisfying \eqref{eq: refining subspace definition}, arbitrary $\var0_0\in \F{U}^0$, and FISTA stepsize choice $(t_n)_{n\in\F N}$. Let $\var0_n$ and $\var1_n$ be generated by Algorithm~\ref{alg: refining FISTA}, then, for any choice of $\var2_n\in\F{U}^n$ and $N\in\F N$ we have
		}
		\begin{equation}\label{eq: FISTA inequality}\Copy{thm:eq: mini FISTA convergence}{
			t_N^2\op{E}_0(\var0_N) + \sum_{n=1}^{N-1}\rho_n \op{E}_0(\var0_n) + \frac{\norm{\var1_N-\var2_N}^2}{2} \leq \frac{\norm{\var0_0-\var2_0}^2-\norm{\var2_0}^2+\norm{\var2_N}^2}{2}
			+ \sum^N_{n=1} t_n\op{E}_0(\var2_n) + \IP{\var1_{n-1}}{\var2_{n-1}-\var2_n}.
		}\end{equation}
	\end{theorem}
	\noindent The proof is given in Theorem~\ref{app:thm: mini FISTA convergence}. This result is the key approximation for showing convergence of FISTA with changing subsets. In the classical setting, we have $\F{U}^n=\F H$, $\var2_n=\var2_0\in\argmin_{\var0\in\F H}\op{E}(\var0)$ and the extra terms on the right-hand side collapse to 0.

	If there exists a minimiser $\var0^*\in\argmin_{\var0\in\F H}\op{E}(\var0)$, then the natural choice in \eqref{eq: FISTA inequality} is $\var2_n=\tens{\Pi}_n\var0^*$ for some projection $\tens{\Pi}_n\colon\F H\to\F{U}^n$, however, there are simple counter-examples which give $\op{E}(\tens{\Pi}_n\var0^*)=\infty$ and so this inequality becomes useless. For example, if $\op{f}(\var0) = \norm{\var0}_{L^2([0,1])}^2$, $\op{g}$ is the indicator on the set $\F D=\{\var0\in L^1([0,1])\st \var0(x) \geq x\}$, and $\tens{\Pi}_n$ is the $L^2$ projection onto a set of piecewise constant functions, then $\var0^*=x\mapsto x$. On the other hand, suppose one of the pixels of the discretisation is $[x_0-\meshsize,x_0+\meshsize]$, then 
	$$\tens{\Pi}_n\var0^*\left(x_0+\tfrac\meshsize2\right) = \argmin_{\vars5\in\R}\int_{x_0-\meshsize}^{x_0+\meshsize}(\var0^*(x)-\vars5)^2\diff x = \argmin_{\vars5\in\R}\int_{x_0-\meshsize}^{x_0+\meshsize}(x-\vars5)^2\diff x = x_0 < x_0+\tfrac\meshsize2.$$
	In particular $\tens{\Pi}_n\var0^* \notin\F D$ therefore $\op{E}(\tens{\Pi}_n\var0^*) = \infty$. The choice $\var2_n=\argmin_{\var0\in\F{U}^n}\op{E}(\var0)$ is much more robust and allows us to apply Algorithm~\ref{alg: refining FISTA} more broadly. The penalty for this flexibility is a more complicated analysis; each time the subset changes, because $\var1_n\in\F{U}_n$, the system receives a `shock' proportional to $\norm{\var1_n}\norm{\var2_{n}-\tens{\Pi}_n\var2_{n+1}}$.

@@

	\subsection{Refinement without overfitting}
	The aim of Step~\ref{step: step 4 sufficiently fast} is to show that $n$ iterations of Algorithm~\ref{alg: refining FISTA} is no slower (up to a constant factor) than $n$ iterations of classical FISTA on the space $\F{U}^n$. In other words, we would like to ensure that
	\begin{equation}
		\op{E}(\var0_n) - \min_{\var0\in\F{U}^n}\op{E}(\var0) \lesssim \frac{\norm{\var0_0-\edit{\argmin_{\var0\in\F{U}^n}\op{E}(\var0)}{\tilde{\var2}_k}}^2}{n^2}
	\end{equation}
	uniformly for $n\in[n_k,n_{k+1})$. If this condition is not satisfied, then it indicates that computational effort has been wasted by a poor choice of subsets. This can be interpreted as an overfitting to the discretisation of $\op{E}|_{\F{U}^n}$ rather than the desired function $\op{E}|_{\F H}$. Combining the assumptions given by Definition~\ref{def: exp subspaces} and the result of Lemma~\ref{thm: mini exponential FISTA convergence}, the following lemma proves the convergence of Algorithm~\ref{alg: refining FISTA} provided that the refinement times $n_k$ are sufficiently small (i.e. $\F{U}^n$ grows sufficiently quickly).

@@

	\subsection{Convergence rate}
	In Lemma~\ref{thm: sufficiently fast} we show that $\op{E}(\var0_n)$ converges at a rate depending on $k$ and $n$, so long as $k$ grows sufficiently quickly. On the other hand, as $k$ grows, the rate becomes worse and so we need to also put a lower limit on the growth of $n_k$. The following lemma completes Step~\ref{step: step 5 sufficiently slow} by computing the global convergence rate of $\op{E}_0(\var0_n)$ when $k$ grows at the minimum rate which is consistent with Lemma~\ref{thm: sufficiently fast}.

@@

	Inserting the computed rates into Theorem~\ref{thm: exponential FISTA convergence} or Theorem~\ref{thm: stronger exponential FISTA convergence} gives the guaranteed convergence rate
	\begin{equation}\label{eq: Lasso energy rate}
		\kappa = \frac{\log \aU^2}{\log \aE + \log \aU^2} = \frac{d}{1+d} \quad \implies \quad \op{E}_0(\var0_n) \lesssim n^{-2(1-\kappa)} = n^{-\frac{2}{1+d}}.
	\end{equation}
	This rate can be used to infer the required resolution at each iteration, in particular on iteration $n$ with $n^2\simeq (\aE\aU^2)^k$ we expect the resolution to be
	\begin{equation}\label{eq: Lasso resolution rate}
		2^{-k} = \left(\aE\aU^2\right)^{\frac{k}{1+d}} \simeq n^{-\frac{2}{1+d}}.
	\end{equation}

>>>>

	\section{Definitions and notation}\label{sec: prelims}
	We consider optimisation of \eqref{eq: min F = fD+fC} over a Hilbert space $(\F H,\IP\cdot\cdot,\norm\cdot)$. In the more analytical section (Sections~\ref{sec: recipe} and \ref{sec: FISTA convergence}) it will be more convenient to use the translated energy
	\begin{equation}
		\op{E}_0\colon\F H\to \R,\qquad \op{E}_0(\var0) \coloneqq \op{E}(\var0)-\Emin[\tilde{\var0}]
	\end{equation}
	so that $\inf_{\var0\in\F{H}}\op{E}_0(\var0) = 0$, although access to this function is not assumed for numerical examples.

@@

	\begin{algorithm}\caption{Refining subset FISTA}\label{alg: refining FISTA}
		\centering
		\begin{algorithmic}[1]
			\State Choose $(\F{U}^n)_{n\in\F N}$, $\var0_0\in \F{U}^0$ and some FISTA stepsize choice $(t_n)_{n\in\F N}$
			\State $\var1_0\gets \var0_0, n\gets0$
			\Repeat
			\State $\bar{\var0}_n \gets (1-\tfrac{1}{t_n})\var0_n + \tfrac{1}{t_n}\var1_n$
			\State $\displaystyle \var0_{n+1} \gets \argmin_{\var0\in \F{U}^{n+1}} \tfrac12\norm{\var0-\bar{\var0}_n+\nabla \op{f}(\bar{\var0}_n)}^2+\op{g}(\var0)$ \Comment{Only modification, $\F{U}^{n+1}\subset \F{U}$}
			\State $\var1_{n+1} \gets (1-t_n)\var0_n + t_n\var0_{n+1}$
			\State $n\gets n+1$
			\Until{some stopping criterion is met}
		\end{algorithmic}
	\end{algorithm}

	\section{General proof recipe}\label{sec: recipe}
	In this section we give an intuitive outline of the full proof for convergence of Algorithm~\ref{alg: refining FISTA} before giving formal theorems and proofs in the next section. First we recall the classical FISTA convergence guarantee given by \cite[Thm 3.1]{Chambolle2015}\edit{:}{; if there exists $\var0^*\in \argmin_{\var0\in\F H}\op{E}_0(\var0)$, then}
	\begin{equation}\label{eq: classical FISTA}
		t_N^2\op{E}_0(\var0_N) + \sum^{N-1}_{n=1}\rho_n\op{E}_0(\var0_n) + \tfrac12\norm{\var1_N-\var0^*}^2 \leq \tfrac12\norm{\var0_0-\var0^*}^2
	\end{equation}
	for \edit{some $\var0^*\in \argmin_{\var0\in\F H}\op{E}_0(\var0)$, and}{any} FISTA stepsize choice $t_N\simeq N$ such that $\rho_n\geq0$.

@@

	\paragraph{Step 2: Quantifying the scaling properties}
	\begin{adaptiveStep}\label{step: step 2 smoothness}
		To show that the extra terms in \eqref{eq: mini FISTA convergence} are small, we need to quantify the approximation properties of $\F{U}^n$. \edit{The idea is that $\F{U}^n$ are locally compact, so that minimisers are achieved, but `growing' so that $\min_{\var0\in\F{U}^n}\op{E}_0(\var0) \to \inf_{\var0\in\F H}\op{E}_0(\var0)$.  A canonical example might be $\F{U}^n=\{\var0\in\F H\st \norm{\var0}\leq n\}$.}{The idea is that there is a sequence $\var2_n\in\F{U}^n$, $n\in\F N$ such that $\norm{\var2_n}$ grows slowly and $\op{E}_0(\var2_n)$ decreases quickly.} To quantify \edit{the}{this} balance\edit{ between a growing Hilbert norm and improving estimate of $\op{E}$}{}, we introduce a secondary sequence $n_0 < n_1 < \ldots$ and constants $\aU,\aE\geq1$ such that for each $k\in\F N$
		\begin{equation}\label{eq: first regularity}
			n\leq n_k \implies \norm{\var2_n} \lesssim \aU^k, \qquad n\geq n_k \implies \op{E}_0(\var2_n) \lesssim \aE^{-k}.
		\end{equation}
		\edit{The expanding balls example would become}{A canonical example would be} $\F{U}^n=\{\var0\in\F H\st \norm{\var0}\leq \aU^k\}$ for $n\in[n_k,n_{k+1})$, then $\aE$ reflects the smoothness of $\op{E}_0$. The choice of exponential scaling \edit{is more}{is introduced to improve stability of Algorithm~\ref{alg: refining FISTA}. It is} natural if we consider the $\F{U}^n$ to be the subspace of functions discretised on a uniform mesh. If that mesh is sequentially refined, then the resolution of the mesh will be of order $h^k$ after $k$ refinements and for some $h<1$. The integer $n_k$ is then the time at which the mesh has refined $k$ times. \edit{The value $\aU$ reflects the smoothness of the chosen basis, and $\aE$ is a measure of efficiency with respect to the structure of $\op{E}$. If an efficient basis is used, then $\aE$ is large.}{} The trade-off between $\aE$ and $\aU$ dictates the final convergence rate of the algorithm. If $\aU>1$, then we cannot guarantee the original $n^{-2}$ rate of convergence.
	\end{adaptiveStep}

@@

		Spatial adaptivity is robust, so long as the scaling properties of Step~\ref{step: step 2 smoothness} are satisfied. The only other constraint is to ensure that the partial telescoping of Step~\ref{step: step 3 convergence bound} still holds.%
		}{%
		The main challenge for adaptive choice of $\F{U}^n$ is to guarantee \eqref{eq: first regularity} from Step~\ref{step: step 2 smoothness} using a posteriori estimates. Combined with the partial telescoping requirement in Step~\ref{step: step 3 convergence bound}, a natural choice is $\var2_n=\var0_{n_k-1}$ for $n\in[n_k,n_{k+1})$, i.e. the value of $n_k$ is chosen to be $n+1$ once the iterate $\var0_n$ is observed.
		}
		Theorem~\ref{thm: stronger exponential FISTA convergence} shows that a sufficient condition is
		$$\var0_{n_k-1}\in\F{U}^{n_k}\cap\F{U}^{n_k+1}\cap\ldots\cap\F{U}^{n_{k+1}-1}\edit{}{, \qquad \norm{\var0_{n_k-1}} \lesssim \aU^k, \quad\text{and}\quad \op{E}_0(\var0_{n_k-1}) \lesssim \aE^{-k}}.$$
		Convergence is most stable if the approximation spaces $\F{U}^n$ satisfy a monotone inclusion, breaking the monotonicity requires more care.
		\edit{Adaptively choosing the time of refinement is more challenging due to the non-descent property of FISTA. The idea is that $\F{U}^n$ should be refined as soon as the value $\op{E}_0(\var0_n)$ is sufficiently close to $\op{E}_0(\var2_n)$. This is analysed in Theorem~\ref{thm: stronger exponential FISTA convergence} (sufficient conditions in Lemma~\ref{thm: practical refinement criteria}) which shows convergence results of the form}{The only non-trivial property to verify is the energy gap  $\op{E}_0(\var0_n) = \op{E}(\var0_n)-\Emin$. Lemma~\ref{thm: practical refinement criteria} proposes some sufficient conditions to guarantee the same overall rate of convergence as in Step~\ref{step: step 5 sufficiently slow},}
		$$ \min_{n\leq N}\op{E}_0(\var0_n) \lesssim \frac{N^{2\kappa}}{N^2}$$
		for all $N\in\F N$, with the same $\kappa\in[0,1)$ from Step~\ref{step: step 5 sufficiently slow}. The penalty for accelerating the change of discretisation is a potential loss of stability or monotonicity in $\op{E}_0(\var0_n)$, although this behaviour has not been seen in numerical experiments.
	\end{adaptiveStep}

@@

	\subsubsection{Single iterations}
	We first wish to understand a single iteration of Algorithm~\ref{alg: refining FISTA}. This is done through the following two lemmas.
	\begin{lemma}[equivalent to {\cite[Lemma 3.1]{Chambolle2015}}]\label{thm: descent lemma}
		\Copy{thm: descent lemma}{
			Suppose $\nabla \op{f}$ is 1-Lipschitz, for any $\bar{\var0}\in \F H$ define
			$$ \var0 \coloneqq \argmin_{\var0\in \F{U}^{n}} \tfrac12\norm{\var0-\bar{\var0}+\nabla \op{f}(\bar{\var0})}^2+\op{g}(\var0).$$
			Then, for all $\var2\in\F{U}^n$, we have
			$$\op{E}_0(\var0) + \tfrac12\norm{\var0-\var2}^2 \leq \op{E}_0(\var2) + \tfrac12\norm{\bar{\var0}-\var2}^2.$$

@@

	\begin{lemma}[{\cite[(17)]{Chambolle2015}, \cite[Lemma 4.1]{Beck2009}}]\label{thm: one step FISTA}
		\Copy{thm: one step FISTA}{
			Let $\var2_n\in\F{U}^n$ be chosen arbitrarily and $\var0_n$/$\var1_n$ be generated by Algorithm~\ref{alg: refining FISTA} for all $n\in\F N$.
			For all $n\in\F N$, it holds that}
		\begin{equation}\Copy{thm:eq: one step FISTA}{
			t_{n}^2(\op{E}_0(\var0_{n}) - \op{E}_0(\var2_n)) - (t_{n}^2-t_{n})(\op{E}_0(\var0_{n-1})-\op{E}_0(\var2_n)) \leq \tfrac1{2}\left[\norm{\var1_{n-1}}^2-\norm{\var1_{n}}^2\right] + \IP{\var1_{n}-\var1_{n-1}}{\var2_n}.}\label{eq: one-step FISTA}
		\end{equation}
	\end{lemma}
	\noindent The proof is given in Theorem~\ref{app:thm: one step FISTA} and is a result of the convexity of $\op{E}_0$ and $\F{U}_n$ for a well chosen $\var2$ in Lemma~\ref{thm: descent lemma}.

@@

	\begin{theorem}[analogous to {\cite[Thm 3.2]{Chambolle2015}, \cite[Thm 4.1]{Beck2009}}]\label{thm: mini FISTA convergence}
		\Copy{thm: mini FISTA convergence}{
			Fix a sequence of subsets $(\F{U}^n)_{n\in\F N}$ satisfying \eqref{eq: refining subspace definition}, arbitrary $\var0_0\in \F{U}^0$, and FISTA stepsize choice $(t_n)_{n\in\F N}$. Let $\var0_n$ and $\var1_n$ be generated by Algorithm~\ref{alg: refining FISTA}, then, for any choice of $\var2_n\in\F{U}^n$ and $N\in\F N$ we have
		}
		\begin{equation}\label{eq: FISTA inequality}\Copy{thm:eq: mini FISTA convergence}{
			t_N^2\op{E}_0(\var0_N) + \sum_{n=1}^{N-1}\rho_n \op{E}_0(\var0_n) + \frac{\norm{\var1_N-\var2_N}^2}{2} \leq \frac{\norm{\var0_0-\var2_0}^2-\norm{\var2_0}^2+\norm{\var2_N}^2}{2}
			+ \sum^N_{n=1} t_n\op{E}_0(\var2_n) + \IP{\var1_{n-1}}{\var2_{n-1}-\var2_n}.
		}\end{equation}
	\end{theorem}
	\noindent The proof is given in Theorem~\ref{app:thm: mini FISTA convergence}. This result is the key approximation for showing convergence of FISTA with changing subsets. In the classical setting, we have $\F{U}^n=\F H$, $\var2_n=\var2_0\in\argmin_{\var0\in\F H}\op{E}_0(\var0)$ and the extra terms on the right-hand side collapse to 0.

	If there exists a minimiser $\var0^*\in\argmin_{\var0\in\F H}\op{E}_0(\var0)$, then the natural choice in \eqref{eq: FISTA inequality} is $\var2_n=\tens{\Pi}_n\var0^*$ for some projection $\tens{\Pi}_n\colon\F H\to\F{U}^n$, however, there are simple counter-examples which give $\op{E}_0(\tens{\Pi}_n\var0^*)=\infty$ and so this inequality becomes useless. For example, if $\op{f}(\var0) = \norm{\var0}_{L^2([0,1])}^2$, $\op{g}$ is the indicator on the set $\F D=\{\var0\in L^1([0,1])\st \var0(x) \geq x\}$, and $\tens{\Pi}_n$ is the $L^2$ projection onto a set of piecewise constant functions, then $\var0^*=x\mapsto x$. On the other hand, suppose one of the pixels of the discretisation is $[x_0-\meshsize,x_0+\meshsize]$, then 
	$$\tens{\Pi}_n\var0^*\left(x_0+\tfrac\meshsize2\right) = \argmin_{\vars5\in\R}\int_{x_0-\meshsize}^{x_0+\meshsize}(\var0^*(x)-\vars5)^2\diff x = \argmin_{\vars5\in\R}\int_{x_0-\meshsize}^{x_0+\meshsize}(x-\vars5)^2\diff x = x_0 < x_0+\tfrac\meshsize2.$$
	In particular $\tens{\Pi}_n\var0^* \notin\F D$ therefore $\op{E}_0(\tens{\Pi}_n\var0^*) = \infty$. The choice $\var2_n=\argmin_{\var0\in\F{U}^n}\op{E}_0(\var0)$ is much more robust and allows us to apply Algorithm~\ref{alg: refining FISTA} more broadly. The penalty for this flexibility is a more complicated analysis; each time the subset changes, because $\var1_n\in\F{U}_n$, the system receives a `shock' proportional to $\norm{\var1_n}\norm{\var2_{n}-\tens{\Pi}_n\var2_{n+1}}$.

@@

	\subsection{Refinement without overfitting}
	The aim of Step~\ref{step: step 4 sufficiently fast} is to show that $n$ iterations of Algorithm~\ref{alg: refining FISTA} is no slower (up to a constant factor) than $n$ iterations of classical FISTA on the space $\F{U}^n$. In other words, we would like to ensure that
	\begin{equation}
		\op{E}_0(\var0_n) = \op{E}(\var0_n) - \min_{\var0\in\F{U}^n}\op{E}(\var0) \lesssim \frac{\norm{\var0_0-\edit{\argmin_{\var0\in\F{U}^n}\op{E}(\var0)}{\tilde{\var2}_k}}^2}{n^2}
	\end{equation}
	uniformly for $n\in[n_k,n_{k+1})$. If this condition is not satisfied, then it indicates that computational effort has been wasted by a poor choice of subsets. This can be interpreted as an overfitting to the discretisation of $\op{E}_0|_{\F{U}^n}$ rather than the desired function $\op{E}_0|_{\F H}$. Combining the assumptions given by Definition~\ref{def: exp subspaces} and the result of Lemma~\ref{thm: mini exponential FISTA convergence}, the following lemma proves the convergence of Algorithm~\ref{alg: refining FISTA} provided that the refinement times $n_k$ are sufficiently small (i.e. $\F{U}^n$ grows sufficiently quickly).

@@

	\subsection{Convergence rate}
	In Lemma~\ref{thm: sufficiently fast} we show that $\op{E}_0(\var0_n)$ converges at a rate depending on $k$ and $n$, so long as $k$ grows sufficiently quickly. On the other hand, as $k$ grows, the rate becomes worse and so we need to also put a lower limit on the growth of $n_k$. The following lemma completes Step~\ref{step: step 5 sufficiently slow} by computing the global convergence rate of $\op{E}_0(\var0_n)$ when $k$ grows at the minimum rate which is consistent with Lemma~\ref{thm: sufficiently fast}.

@@

	Inserting the computed rates into Theorem~\ref{thm: exponential FISTA convergence} or Theorem~\ref{thm: stronger exponential FISTA convergence} gives the guaranteed convergence rate
	\begin{equation}\label{eq: Lasso energy rate}
		\kappa = \frac{\log \aU^2}{\log \aE + \log \aU^2} = \frac{d}{1+d} \quad \implies \quad \op{E}(\var0_n)-\Emin \lesssim n^{-2(1-\kappa)} = n^{-\frac{2}{1+d}}.
	\end{equation}
	This rate can be used to infer the required resolution at each iteration, in particular on iteration $n$ with $n^2\simeq (\aE\aU^2)^k$ we expect the resolution to be
	\begin{equation}\label{eq: Lasso resolution rate}
		2^{-k} = \left(\aE\aU^2\right)^{\frac{k}{1+d}} \simeq n^{-\frac{2}{1+d}}.
	\end{equation}
